---
title: "baseballCurrentYear"
author: "Ajoni Hopkins"
date: "2024-04-27"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(dev = 'pdf')
```

```{r, echo=FALSE, message=FALSE}
# Loading Libraries ----
library(tidyverse)
#library(YFAR)
library(knitr)
#library(tinytex)
library(kableExtra)
library(RColorBrewer)
```

```{r loading-datasets, echo=FALSE}
# Loading Datasets  ----

load("Output Data/teams.RDS")
load("Output Data/allWeeks.RDS")
load("Output Data/stat_cats.RDS")

```

This report will export visuals about our Fantasy Baseball League. It will first export tables/visuals about the entire league, and afterwards, visuals about a particular team. Most of the visuals will filtered to show data from a specific week, team, and sometimes a particular statistical category.

In the section below, you will see the variables selected for this report.

## Variables for the Report
```{r notes, echo = FALSE}
# Notes for future dev
# create an r script that makes my static fifles, that runs weekly monday in the AM.
# the shiny taps into my local files that get updated!

# will creating these filters, only update ethe visuals that have cats that change? or will it re-run the entire report? we want the former
# create a select all field for teams to include in the all team analysis.




# map linetypes to specific values
```

```{r variables-for-report, echo = TRUE}
# Team to analyze 
teamAnalyze <- "Aj's Awesome Team" # select 1 only

# week range for data viz
# number of completed weeks
completedWeeks <- allWeeks %>%
  filter(team_remaining_games == 0) %>% # only getting info for completed weeks
  select(team_stats_week) %>%
  distinct(team_stats_week) %>%
  as_vector()
  
# week to analyze
weekAnalyze <- 6 # select one only

#weekRangeAnalyze <- c(3,4) # this will be a filter button, select all that apply
weekRangeAnalyze <- completedWeeks

# selected stats for dataviz
battingStatSelected <- 'RBI' # statsListHitting
pitchingStatSelected <- 'K' #statsListPitching

# Visualization Options ----
statsList <- c('R', 'HR', 'RBI', 'SB', 'OPS',
               'W', 'K', 'ERA', 'WHIP', 'SVH')

statsListHitting <- c('R', 'HR', 'RBI', 'SB', 'OPS')

statsListPitching <- c('W', 'K', 'ERA', 'WHIP', 'SVH')

```

```{r data-manipulation, echo = FALSE, message=FALSE}
# Data Manipulation ----


# creating colors for the data viz
myColors <- brewer.pal(12, "Paired")
names(myColors) <- levels(allWeeks$team_name)
colScale <- scale_colour_manual(name = 'team_name', values = myColors)
colScaleFill <- scale_fill_manual(name = 'team_name', values = myColors)

# creating line type for the data viz
myLines <- c('solid', 'twodash','solid', 'twodash','solid', 'twodash',
             'solid', 'twodash','solid', 'twodash','solid', 'twodash')
names(myLines) <- levels(allWeeks$team_name)
lineType <- scale_linetype_manual(name = 'team_name', values = myLines)

# getting data for every team, for every completed week
allWeekStats <- allWeeks %>%
  filter(team_remaining_games == 0) %>% # only getting info for completed weeks
  select(team_stats_week, team_key, team_id, team_name, team_stats) %>%
  unnest(team_stats)

allWeekStats2 <- left_join(allWeekStats, stat_cats, by = 'stat_id') %>%
  filter(display_name != 'H/AB') %>%
  mutate(value = as.numeric(value)) %>%
  select(-stat_id)

# ranking results by week and category. creating z scores
allWeekStats3 <- allWeekStats2 %>%
  group_by(team_stats_week, display_name) %>%
  mutate(Rank = rank(-value),
         zScore = ifelse(display_name %in% c('ERA', 'WHIP'),
                         -(value - mean(value))/sd(value),
                         (value - mean(value))/sd(value)))

# getting max range of zScore for data viz
zScoreRange <- allWeekStats3 %>%
  ungroup() %>%
  select(zScore) %>%
  mutate(zScore = round(abs(zScore), 0)) %>%
  filter(zScore == max(zScore)) %>%
  arrange(desc(zScore)) %>%
  distinct(zScore) %>%
  as.numeric()

# Testing Function -----------------------------------------------------
# allWeekStats3 %>%
#     ungroup() %>%
#     filter(team_stats_week == 6) %>%
#     select(-c(Rank, zScore, team_stats_week, team_key, position_type)) %>%
#     pivot_wider(names_from = display_name, values_from = value) %>%
#     mutate(Week = paste('Week', 7), .before = team_name) %>%
#     relocate(IP,
#              R, HR, RBI, SB, OPS,
#              W, K, ERA, WHIP, `SV+H`, .after= team_name) %>%
#     rename(SVH = `SV+H`)

# Testing Function End -------------------------------------------------


# creating function for a results table by week for all source stats! ----
actualValuesWeek <- function(weekNum){
  actualValues <- allWeekStats3 %>%
    ungroup() %>%
    filter(team_stats_week == weekNum) %>%
    select(-c(Rank, zScore, team_stats_week, team_key, position_type)) %>%
    pivot_wider(names_from = display_name, values_from = value) %>%
    mutate(Week = paste('Week', weekNum), .before = team_name) %>%
    relocate(IP,
             R, HR, RBI, SB, OPS,
             W, K, ERA, WHIP, `SV+H`, .after= team_name) %>%
    rename(SVH = `SV+H`)
  
  actualValues
  
}

# creating function for a results table by week for rankings! ----
rankingsWeek <- function(weekNum){
  
  rankings <- allWeekStats3 %>%
    ungroup() %>%
    filter(team_stats_week == weekNum) %>%
    select(-c(value, zScore, team_stats_week, team_key, position_type)) %>%
    spread(key = display_name, value = Rank) %>%
    mutate(Week = paste('Week', weekNum), .before = team_name) %>%
    relocate(IP,
             R, HR, RBI, SB, OPS,
             W, K, ERA, WHIP, `SV+H`, .after= team_name) %>%
    rename(SVH = `SV+H`) 
  
  rankings
}
  
# creating function for a results table by week for z scores! ----
zScoreWeek <- function(weekNum){

  zScore <- allWeekStats3 %>%
    ungroup() %>%
    filter(team_stats_week == weekNum) %>%
    select(-c(Rank, value, team_stats_week, team_key, position_type)) %>%
    spread(key = display_name, value = zScore) %>%
    mutate(Week = paste('Week', weekNum), .before = team_name) %>%
    relocate(IP,
             R, HR, RBI, SB, OPS,
             W, K, ERA, WHIP, `SV+H`, .after= team_name) %>%
    rename(SVH = `SV+H`)

  zScore
}

# creating function for a results table by week for total z scores by hitting/pitching! ----
zScoreSumsWeek <- function(weekNum){
  
  zScoreSumsCat <- allWeekStats3 %>%
    ungroup() %>%
    filter(team_stats_week == weekNum) %>%
    mutate(Week = paste('Week', weekNum),
           weekNumber = weekNum) %>%
    group_by(Week, weekNumber, team_name, position_type) %>%
    summarise(zScoreSum = sum(zScore)) %>%
      spread(key = team_name, value = zScoreSum)
  
  
  zScoreSumsCat2 <- allWeekStats3 %>%
    ungroup() %>%
    filter(team_stats_week == weekNum) %>%
    mutate(Week = paste('Week', weekNum),
           weekNumber = weekNum) %>%
    group_by(Week, weekNumber, team_name) %>%
    summarise(zScoreSum = sum(zScore)) %>%
      spread(key = team_name, value = zScoreSum) %>%
      mutate(position_type = 'Total')
  
  zScoreSumsCat3 <- rbind(zScoreSumsCat, zScoreSumsCat2)
  
}


# creating function for a results table by week for total z scores by hitting/pitching! ----
zScoreSums <- tibble()

for (c in weekRangeAnalyze ){
  zScoreSumsadd <- zScoreSumsWeek(c)
  
  zScoreSums <- rbind(zScoreSums, zScoreSumsadd)
}

zScoreSums2 <- zScoreSums %>%
  gather(key = 'team_name', value = 'zScoreSum', 4:15)

# Creating the tables functions ----
actualValues <- actualValuesWeek(weekAnalyze)
rankings <- rankingsWeek(weekAnalyze)
zScore <- zScoreWeek(weekAnalyze)
zScoreSums <- zScoreSumsWeek(weekAnalyze)
```
  
  
  
## Full League Viz

```{r full-league-tables, include = TRUE, echo=FALSE, fig.align = 'center'}

# 1 Week Data Viz
kable(actualValues, caption = "Actual Results by Team for Selected Week")

kable(rankings, caption = "Rankings by Team for Selected Week [1 High]")

kable(zScore, digits = 4, caption = "Z-Scores by Team for Selected Week")

kable(zScoreSums, digits = 4, caption = "Sum of Z-Scores by Stat Category and Team for Selected Week Range") %>%
  kable_styling(bootstrap_options = c('striped', 'condensed', 'responsive'),
                full_width = F,
                position = 'center') %>%
  row_spec(0, font_size=9)

```


```{r full-league-viz-1, include = TRUE, echo=FALSE, message=FALSE, fig.width = 10, fig.align = 'center'}

# Total ZScore by Team
zScoreSums2 %>%
  filter(position_type == 'Total',
         weekNumber %in% weekRangeAnalyze) %>%
  group_by(team_name) %>%
  summarise(zScoreSum = sum(zScoreSum)) %>%
  ggplot(aes(x = team_name, y = zScoreSum, fill = team_name)) +
    geom_col() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    labs(title = 'Total Z-Score for Selected Week Range') +
    geom_hline(yintercept = 0, color = 'dark gray') +
    colScaleFill
    
```

```{r full-league-viz-2, include = TRUE, echo=FALSE, message=FALSE, out.extra='style="max-width:none; width:90vw; margin-left:calc(50% - 40vw);"', fig.height = 2.7}

# z score for each category!
zScoreSums2 %>%
  filter(position_type != 'Total',
         weekNumber %in% weekRangeAnalyze) %>%
  group_by(team_name, position_type) %>%
  summarise(zScoreSum = sum(zScoreSum)) %>%
  ggplot(aes(x = team_name, y = zScoreSum, fill = position_type)) +
    geom_col(position = 'dodge') +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    labs(title = 'Total Z-Score for Selected Week Range by Stat Type') +
    theme(text=element_text(size=6),
          legend.key.size = unit(.5, 'cm')) +
    geom_hline(yintercept = 0, color = 'dark gray')

#axis.text=element_text(size=16),
#          legend.text=element_text(size=16),
#          axis.title = element_text(size=16)

```

```{r full-league-viz-3, include = TRUE, echo=FALSE, message=FALSE, fig.width = 10, fig.align = 'center'}
# Compare All Teams, 1 Stat, Multiple Weeks, Hitting
allWeekStats3 %>%
  filter(display_name == battingStatSelected,
         team_stats_week %in% weekRangeAnalyze) %>%
  ggplot(aes(x = team_stats_week, y = zScore, color = team_name, group = team_name, linetype = team_name)) +
    geom_line(linewidth = 1) +
    geom_point() +
    labs(title = paste('Z-Score for Selected Week Range for', battingStatSelected, sep = ' ')) +
    xlab(label = 'Week Number') +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray') +
    colScale +
    lineType

# Compare All Teams, 1 Stat, Zscores Meaned, Hitting
allWeekStats3 %>%
  filter(display_name == battingStatSelected,
         team_stats_week %in% weekRangeAnalyze) %>%
  group_by(team_name) %>%
  summarise(zScore = mean(zScore, na.rm = TRUE)) %>%
  ggplot(aes(x = team_name, y = zScore, fill = team_name)) +
    geom_col() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    labs(title = paste('Mean Z-Score for Selected Week Range for', battingStatSelected, sep = ' ')) +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray') +
    colScaleFill

# Compare All Teams, 1 Stat,Multiple Weeks, Pitching
allWeekStats3 %>%
  filter(display_name == pitchingStatSelected,
         team_stats_week %in% weekRangeAnalyze) %>%
  ggplot(aes(x = team_stats_week, y = zScore, color = team_name, group = team_name, linetype = team_name)) +
    geom_line(linewidth = 1) +
    geom_point() +
    labs(title = paste('Z-Score for Selected Week Range for', pitchingStatSelected, sep = ' ')) +
    xlab(label = 'Week Number') +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray') +
    colScale +
    lineType

# Compare All Teams, 1 Stat, Zscores Mean, Pitching
allWeekStats3 %>%
  filter(display_name == pitchingStatSelected,
         team_stats_week %in% weekRangeAnalyze) %>%
  group_by(team_name) %>%
  summarise(zScore = mean(zScore, na.rm = TRUE)) %>%
  ggplot(aes(x = team_name, y = zScore, fill = team_name)) +
    geom_col() +
    scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
    labs(title = paste('Mean Z-Score for Selected Week Range for', pitchingStatSelected, sep = ' ')) +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray') +
    colScaleFill

```

## Team Specific Viz

```{r team-specific-viz-HP, echo = FALSE}
# Both Hitting and Pitching ----
# zScore, Compare stats 1 team
zScoreSums2 %>%
  filter(team_name == teamAnalyze,
         weekNumber %in% weekRangeAnalyze) %>%
  ggplot(aes(x = weekNumber, y = zScoreSum, color = position_type, group = position_type)) +
    geom_line() +
    geom_point() +
    labs(title = paste('Z-Score Sum for Selected Week Range by Stat Category for', teamAnalyze, sep = ' ')) +
    geom_hline(yintercept = 0, color = 'dark gray')

```

```{r team-specific-viz-hitting, echo=FALSE}
# Hitting Categories ----
# 1 team, all selected weeks
allWeekStats3 %>%
  filter(position_type == 'B', team_name == teamAnalyze,
         team_stats_week %in% weekRangeAnalyze) %>%
  ggplot(aes(x = as.numeric(team_stats_week), y = zScore, color = display_name)) +
    geom_line() +
    geom_point() +
    labs(title = paste('Z-Score by Week and Hitting Category for', teamAnalyze, sep = ' ')) +
    xlab(label = 'Week Number') +
    scale_color_discrete(name = "Category") +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray')

# Mean z-score by category, all selected weeks
allWeekStats3 %>%
  filter(position_type == 'B', team_name == teamAnalyze,
         team_stats_week %in% weekRangeAnalyze) %>%
  group_by(display_name) %>%
  summarise(zScore = mean(zScore, na.rm = TRUE)) %>%
  ggplot(aes(x = display_name, y = zScore)) +
    geom_col() +
    labs(title = paste('Mean Z-Score for Selected Week Range by Hitting Category for', teamAnalyze, sep = ' ')) +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray')

```

```{r team-specific-viz-pitching, echo=FALSE}
# Pitching Categories ----
# 1 team, all selected weeks
allWeekStats3 %>%
  filter(position_type == 'P', team_name == teamAnalyze,
         team_stats_week %in% weekRangeAnalyze) %>%
  ggplot(aes(x = as.numeric(team_stats_week), y = zScore, color = display_name)) +
    geom_line() +
    geom_point() +
    labs(title = paste('Z-Score by Week and Pitching Category for', teamAnalyze, sep = ' ')) +
    xlab(label = 'Week Number') +
    scale_color_discrete(name = "Category") +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray')

# Mean z-score by category, all selected weeks
allWeekStats3 %>%
  filter(position_type == 'P', team_name == teamAnalyze,
         team_stats_week %in% weekRangeAnalyze) %>%
  group_by(display_name) %>%
  summarise(zScore = mean(zScore, na.rm = TRUE)) %>%
  ggplot(aes(x = display_name, y = zScore)) +
    geom_col() +
  labs(title = paste('Z-Score for Selected Week Range by Stat Category for', teamAnalyze, sep = ' ')) +
    xlab(label = 'Week Number') +
  scale_color_discrete(name = "Category") +
    ylim(-zScoreRange, zScoreRange) +
    geom_hline(yintercept = 0, color = 'dark gray')

```

